New lower bounds for Schur and weak Schur numbers

### 2021-05-08 article/weakschur.tex
preuve wsf templates

\qquad In this section, we generalize Rowley's constructions for weak Schur numbers \cite{RowleyWS} and give an analogous
for weak Schur numbers of Abbott and Hanson's construction for Schur numbers. By analogy with the previous section,
we then introduce WSF-templates as well as the sequence \(\WS^+(n)\). We find suitable templates and use them to
provide new lower bounds for weak Schur numbers.

@@

\qquad Up to now, there was no equivalent for weak Schur numbers of Abbott and Hanson's construction for Schur numbers
\cite{AbbottHanson}. Here we give a general lower bound for weak Schur numbers as a function of both regular and
weak Schur numbers. The following theorem, inspired by Rowley's inequalities for \(\WS (n+1)\) and \(\WS (n+2)\),
was found and proved by Romain Ageron.

\begin{theorem}
\label{theorem}
Let \((p,q), (n,k) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free
subsets then there exists a partition of \([\![1,p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q]\!]\) into \(n+k\)
weakly sum-free subsets.
\end{theorem}

In particular, if we choose \(q = \WS (n)\) and \(p = S(k)\) in the last theorem, the next corollary follows.

@@

Let \((p, q) \in \mathbb{N}^2\) such that there exists a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets. Let \(a \in \mathbb{N}\) with \(a > q\)
and let's try to build a coloring of \([\![1, ap + q]\!]\) into \(n + k\) weakly sum-free subsets. Let
\(l = a - b - 1\), \(r \in [\![1,q]\!]\) and \(w = a - l - r - 1 = b - r\).

@@

\renewcommand{\arraystretch}{1.7}

\[
\begin{NiceArray}{|*{13}{c|}}[corners=NW,standard-cline,hlines]
\CodeBefore
	\cellcolor{green}{1-6}
	\cellcolor{green}{1-7}
	\cellcolor{cyan}{1-8}
	\cellcolor{green}{1-9}
	\cellcolor{cyan}{1-10}
	\cellcolor{cyan}{1-11}
	\cellcolor{cyan}{1-12}
	\cellcolor{green}{1-13}
	\cellcolor{red}{2-1}
	\cellcolor{red}{2-2}
	\cellcolor{red}{2-3}
	\cellcolor{red}{2-4}
	\cellcolor{red}{2-5}
	\cellcolor{red}{2-6}
	\cellcolor{red}{2-7}
	\cellcolor{red}{2-8}
	\cellcolor{red}{2-9}
	\cellcolor{cyan}{2-10}
	\cellcolor{cyan}{2-11}
	\cellcolor{cyan}{2-12}
	\cellcolor{green}{2-13}
	\cellcolor{yellow}{3-1}
	\cellcolor{yellow}{3-2}
	\cellcolor{yellow}{3-3}
	\cellcolor{yellow}{3-4}
	\cellcolor{yellow}{3-5}
	\cellcolor{yellow}{3-6}
	\cellcolor{yellow}{3-7}
	\cellcolor{yellow}{3-8}
	\cellcolor{yellow}{3-9}
	\cellcolor{cyan}{3-10}
	\cellcolor{cyan}{3-11}
	\cellcolor{cyan}{3-12}
	\cellcolor{green}{3-13}
	\cellcolor{yellow}{4-1}
	\cellcolor{yellow}{4-2}
	\cellcolor{yellow}{4-3}
	\cellcolor{yellow}{4-4}
	\cellcolor{yellow}{4-5}
	\cellcolor{yellow}{4-6}
	\cellcolor{yellow}{4-7}
	\cellcolor{yellow}{4-8}
	\cellcolor{yellow}{4-9}
	\cellcolor{cyan}{4-10}
	\cellcolor{cyan}{4-11}
	\cellcolor{cyan}{4-12}
	\cellcolor{green}{4-13}
	\cellcolor{red}{5-1}
	\cellcolor{red}{5-2}
	\cellcolor{red}{5-3}
	\cellcolor{red}{5-4}
	\cellcolor{red}{5-5}
	\cellcolor{red}{5-6}
	\cellcolor{red}{5-7}
	\cellcolor{red}{5-8}
	\cellcolor{red}{5-9}
	\cellcolor{cyan}{5-10}
	\cellcolor{cyan}{5-11}
	\cellcolor{cyan}{5-12}
	\cellcolor{green}{5-13}
	\cellcolor{cyan}{6-10}
	\cellcolor{cyan}{6-11}
	\cellcolor{cyan}{6-12}
	\cellcolor{green}{6-13}
\Body
	& & & & & 1 & 2 & ... & r & r + 1 & ... & b - 1 & b \\
	a - l & a - l + 1 & ... & a - 1 & a & a + 1 & ... & a + r - 1 & a + r & a + r + 1 & ... & a + b - 1 & a + b \\
	2 a - l & ... & ... & ... & 2 a & ... & ... & ... & 2 a + r & ... & ... & ... & 2 a + b \\
	... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... \\
	... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... \\
	p a - l & ... & ... & ... & p a & ... & ... & ... & p a + r & ... & ... & ... & p a + b \\
\end{NiceArray}
\]

@@

\begin{itemize}
	\item \underline{\(\lambda(x+y) = \lambda(x) + \lambda(y)\)} \\
	In this case, we use the sum-free property of the coloring of \([\![1,p]\!]\) (in block \(\mathcal{C}\), we only use colors \(1, ..., n\)).

	\item \underline{\(\lambda(x+y) \neq \lambda(x) + \lambda(y)\)} \\
	In this case, we do not have information about the color of \(\lambda(x+y)\). Thereby, we want to have \(x+y \in \mathcal{C}\).
	A simple solution is to limit the horizontal movement, that is if the sum changes line, not to move too far so that it stays in \(\mathcal{C}\).
	There, the maximal displacement to the left (resp. to the right) is \(2l\) (resp. \(2r\)). Not crossing entirely \(\mathcal{C}\) by going to the left
	is then expressed as \(-2l > -a + r\). Likewise, not going to far to the right is expressed as \(2r < a - l\). It can then be written as
	 \(\max(l, r) \leqslant w\).
\end{itemize}

@@

\qquad In this subsection, we introduce WSF-templates and define the objects and prove the results needed
for the general theorem on templates for weak Schur numbers.

\begin{definition}

Let \((a,b) \in (\mathbb{N}^*)^2\) such that \(a>b\). Define
\[ \pi_{a,b}:x \longmapsto (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})(x \mod a)\]
\end{definition}

If there is no confusion on the \(a\) and \(b\) to use, \(\pi_{a, b}\) is denoted by \(\pi\).

\(\pi\) is the projection on the first line mentioned in the intuitive explanation.

\begin{proposition}
Let x \(\in [\![1,b]\!]\) and \(y \in \mathbb{N}^*\) such that \(x+\pi(y) \leqslant a+b\). Then \(\pi(x+y)=x+\pi(y)\).
\end{proposition}

\begin{proof}[\textsc{Proof.}]
Let  \(x\in [\![1,b]\!]\), and  \(y\in \mathbb{N}^*\) such that \(x+\pi(y)\leqslant a+b\).

\noindent if \(x+\pi(y)< a \):

\[
\begin{array}{c c l l}
	\pi(x+y) & = & (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})(x+y \mod a) & \\
	& = & (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y) \mod a) & \text{since } \pi(y)=y \mod a \\
	& = & x+ \pi(y) &
\end{array}
\]

\noindent if \(x+\pi(y)\geqslant a\):
\[
\begin{array}{c c l l}
	\pi(x+y) & = & (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})(x+y \mod a)& \\
	& = & (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y)\mod a) & \text{since } \pi(y)=y \mod a \\
	& = & (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y) - a) & \\
	& = & x+\pi(y) - a +a\mathds{1}_{ [\![0,b]\!]}(x+\pi(y) - a) & \\
	& = & x+\pi(y) - a +a & \text{since } x+\pi(y) - a \in [\![0,b]\!] \\
	& = & x+\pi(y) &
\end{array}
\]
\end{proof}

\begin{proposition}
Let \((x,y)\in (\mathbb{N}^*)^2\). Then \(\pi(\pi(x)+\pi(y))=\pi(x+y)\).
\end{proposition}

\begin{proof}[\textsc{Proof.}]
Let \((x,y)\in (\mathbb{N}^*)^2\).

\[
\begin{array}{c c l}
	\pi(\pi(x)+\pi(y)) & = & (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})(\pi(x)+\pi(y) \mod a)\\
	& = & (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})((Id+a\mathds{1}_{ [\![0,b]\!]})(x \mod a)+(Id+a\mathds{1}_{ [\![0,b]\!]})(y \mod a) \mod a) \\
	& = & (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})(((x \mod a) + (y \mod a)) \mod a)\\
	& = & (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})( x+y \mod a)\\
	& = & \pi(x+y)
\end{array}
\]
\end{proof}

\begin{definition}

Let \((a,b) \in (\mathbb{N}^*)^2\) such that \(a>b\). Define
\[ \lambda_{a,b}:x \longmapsto 1+ \left\lfloor\dfrac{x-b-1}{a}\right\rfloor\]
\end{definition}

@@

\begin{proposition}
Let \((a,b)\in (\mathbb{N}^*)^2\) such that \(a>b\) and let \(x\in \mathbb{N}^*\). Then \(x=a\lambda(x)+\pi(x)-a\).
\end{proposition}

@@

\noindent  if \(x \mod a>b\) then \(a\lambda(x)+\pi(x)-a=a\left\lfloor\dfrac{x}{a}\right\rfloor+x \mod a=x\)

\noindent if \(x \mod a \leqslant b\) then
\(a\lambda(x)+\pi(x)-a=a \left( \left \lfloor \dfrac{x}{a} \right \rfloor - 1 \right)+x \mod a +a=x\)
\end{proof}

\begin{definition}
Let \( (a,n,b) \in (\mathbb{N}^*)^3\), Let \((A_1,...,A_n)\) a partition of  \([\![1, a + b]\!]\).
This partition is said to be a \(b\)-weakly-sum-free template (\(b\)-WSF-template) with width \(a\) and \(n\) colors if

\begin{itemize}
\item \(\forall i \in [\![1, n]\!], \quad A_i\) is weakly-sum-free
\item \(\forall i \in [\![1, n]\!], \quad A_i\backslash [\![1, b]\!]\) is sum-free
\item For \(A_n\) (the special subset)
	\[
	\forall (x,y) \in A_n^2, \,x+y>b+2a \implies x+y-2a\notin A_n
	\]
\item For the others subsets
	\[
	\forall i \in [\![0,n-1]\!], \, \forall(x,y) \in A_i^2, \, x+y>a+b \implies \pi(x+y) \notin A_i
	\]
\end{itemize}
\end{definition}

@@

\begin{definition}
Let \( (n,b) \in (\mathbb{N}^*)^2\). If there exists \(a\) such that there exists a \(b\)-WSF-template with width \(a\) 
and \(n\) colors, define
\[
\WS_b^+(n)=-b+\max \{a \in \mathbb{N}^*/ \text{there exists a } b \text{-WSF-template with width } a \text{ and } n \text{ colors} \}
\]
If no such \(a\) exists, set \(\WS_b^+(n) = 0\).
\end{definition}

@@

\begin{proof}[\textsc{Proof.}]
The lower bound comes from the analogous of Abott and Hanson's construction for weak Schur numbers.
The upper bound comes from the fact that a WSF-template with width \(a\) and \(n\) colors contains a partition of
\([\![1, a]\!]\) into \(n\) sum-free subsets.
\end{proof}

@@

\begin{theorem}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) and \( (p,k) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,p]\!]\) 
into \(k\) sum-free subsets and a \(b\)-WSF-template \((A_1,...,A_n)\) with width \(a\) and \(n\) colors, 
then there exists a partition of \([\![1, b+p \times a]\!]\) into \((k+n-1)\) weakly sum-free subsets.
\end{theorem}

In particular, by setting \(p = S(k)\) and \(a = \WS^+(n)\) in the last theorem, the next corollary follows.

\begin{corollary}
For every \((n,k) \in (\mathbb{N}^*)^2\), define \( b_{max} = \max \{b\in \mathbb{N}^*/ \WS_b^+(n+1) = \WS^+(n+1)\}\).
Then the following inequality holds.
\[\WS(n+k) \geqslant S(k) \WS^+(n+1)+b_{max}\]
\end{corollary}

\begin{remark}
In the SF-template construction for Schur numbers, the additive constant comes from the fact that the special color does
not necessarily appear right at the begining of the repeating pattern. Likewise, \(b_{max}\) can actually be replaced by \\
\[
\max_{b \in \mathbb{N}^*} \left\{\min (A_{n+1} \backslash [\![1, b ]\!]) - 1~|~ \WS_b^+(n+1) = \WS^+(n+1) \right\}
\]
\end{remark}

\begin{proof}[\textsc{Proof.}]
Let \((a,n,b) \in (\mathbb{N}^*)^3\) and \( (p,k) \in (\mathbb{N}^*)^2\). Set \(q = a + b\). \\
We denote by \(g\) the coloring associated to the \(b\)-WSF-template  and \(h\) the one associated 
to the sum-free partition of \([\![1,p]\!]\).

\[
\begin{array}{l l}
	g : [\![1, q]\!] & \longrightarrow [\![1,n]\!] \\
	h : [\![1,p]\!] & \longrightarrow [\![1,k]\!]
\end{array}
\]

Define \( f : [\![1,b+pa]\!] \longrightarrow [\![1,n]\!] \) as follows

\begin{itemize}
	\item if  \(x\leqslant b \text{ (we will note }x \in \mathcal{T}): f(x)=g(x)\)
	\item if \( x \in [\![1,b+pa]\!] \text{ and } \pi(x) \notin A_n (\text{ we will note } x \in \mathcal{C}): f(x)=g(\pi(x))\)
	\item if \( x \in [\![1,b+pa]\!]  \text{ and }\pi(x) \in A_n (\text{ we will note } x \in \mathcal{R}): f(x)=n-1+h(\lambda(x))\)
\end{itemize}

\(f\) is well defined since \(\pi\) is defined for \(x>b\) and \(\forall x \in \![1,b+pa]\!], f(x)\leqslant n+k-1\) because \(h(\lambda(x))\leqslant k\)

We have parted the integers of \([\![1,b+pa]\!]\) in three disjoints subsets \(\mathcal{T},\mathcal{C} \text{ and } \mathcal{R}\).

We have to verify that f induces weakly-sum-free templates\\
\\
\underline{if \((x,y) \in \mathcal{T}^2\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
\\\(x+y<a+b\) since \(b<a\) and \(g(x)=f(x)=f(y)=g(y)\).
\\Hence \(f(x+y)=g(x+y)\neq g(x)=f(x)\)
\\\\
\underline{if \((x,y) \in \mathcal{T} \times \mathcal{C}\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
We distinguish two cases

\begin{itemize}
\item If \(x+\pi(y)\leqslant a+b\)
\\\(g(x)=f(x)=f(y)=g(\pi(y)\). Hence \(g(x)\neq g(x+\pi(y))=g(\pi(x+y))\) (qv previous proposition)
\\if \(g(\pi(x+y))=n, f(x+y)\geqslant n > f(x)\)
\\else, \(f(x+y)=g(\pi(x+y))\neq g(x)=f(x)\)
\item If \(x+\pi(y)> a+b\), \(x+y>a+b\) and by definition of \(g\), \(g(\pi(x+y))\neq g(x)\)
\\if \(g(\pi(x+y))=n, f(x+y)\geqslant n > f(x)\)
\\else, \(f(x+y)=g(\pi(x+y))\neq g(\pi(x))=f(x)\)
\end{itemize}

\underline{if \((x,y) \in \mathcal{T} \times \mathcal{R}\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
\\Then, \(f(x)=f(y)=n\). We distinguish two cases

>>>>

\qquad In this section, we generalize Rowley's constructions for weak Schur numbers \cite{RowleyWS} and give an analogous
for weak Schur numbers of Abbott and Hanson's construction for Schur numbers. By analogy with the previous section,
we then introduce WSF-templates as well as an associated sequence \(\WS^+(n)\). We find templates and use them to
provide new lower bounds for weak Schur numbers.

@@

\qquad Up to now, no equivalent for weak Schur numbers of Abbott and Hanson's construction for Schur numbers
\cite{AbbottHanson} was known. Here we give a general lower bound for weak Schur numbers as a function of both 
regular and weak Schur numbers. The following theorem, inspired by Rowley's inequalities for \(\WS (n+1)\) and 
\(\WS (n+2)\), was found and proved by Romain Ageron.

\begin{theorem}
\label{theorem}
\begin{sloppypar}
Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free
subsets then there exists a partition of \({[\![1,p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q]\!]}\) into \(n+k\)
weakly sum-free subsets.
\end{sloppypar}
\end{theorem}

In particular, by setting \(q = \WS (n)\) and \(p = S(k)\) in the above theorem, one obtains the following corollary.

@@

Let \((p, k), (q, n) \in (\mathbb{N}^*)^2\) such that there exists a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets. Let \(a \in \mathbb{N}\) with \(a > q\)
and let's try to build a coloring of \([\![1, ap + q]\!]\) into \(n + k\) weakly sum-free subsets. Let
\(l = a - b - 1\), \(r \in [\![1,q]\!]\) and \(w = a - l - r - 1 = b - r\).

@@

\renewcommand{\arraystretch}{1.7}
\setlength{\arraycolsep}{4pt}

\[
\begin{NiceArray}{*{13}{c}}[corners=NW,hvlines]
\CodeBefore
	\cellcolor{green}{1-6}
	\cellcolor{green}{1-7}
	\cellcolor{cyan}{1-8}
	\cellcolor{green}{1-9}
	\cellcolor{cyan}{1-10}
	\cellcolor{cyan}{1-11}
	\cellcolor{cyan}{1-12}
	\cellcolor{green}{1-13}
	\cellcolor{red}{2-1}
	\cellcolor{red}{2-2}
	\cellcolor{red}{2-3}
	\cellcolor{red}{2-4}
	\cellcolor{red}{2-5}
	\cellcolor{red}{2-6}
	\cellcolor{red}{2-7}
	\cellcolor{red}{2-8}
	\cellcolor{red}{2-9}
	\cellcolor{cyan}{2-10}
	\cellcolor{cyan}{2-11}
	\cellcolor{cyan}{2-12}
	\cellcolor{green}{2-13}
	\cellcolor{yellow}{3-1}
	\cellcolor{yellow}{3-2}
	\cellcolor{yellow}{3-3}
	\cellcolor{yellow}{3-4}
	\cellcolor{yellow}{3-5}
	\cellcolor{yellow}{3-6}
	\cellcolor{yellow}{3-7}
	\cellcolor{yellow}{3-8}
	\cellcolor{yellow}{3-9}
	\cellcolor{cyan}{3-10}
	\cellcolor{cyan}{3-11}
	\cellcolor{cyan}{3-12}
	\cellcolor{green}{3-13}
	\cellcolor{yellow}{4-1}
	\cellcolor{yellow}{4-2}
	\cellcolor{yellow}{4-3}
	\cellcolor{yellow}{4-4}
	\cellcolor{yellow}{4-5}
	\cellcolor{yellow}{4-6}
	\cellcolor{yellow}{4-7}
	\cellcolor{yellow}{4-8}
	\cellcolor{yellow}{4-9}
	\cellcolor{cyan}{4-10}
	\cellcolor{cyan}{4-11}
	\cellcolor{cyan}{4-12}
	\cellcolor{green}{4-13}
	\cellcolor{red}{5-1}
	\cellcolor{red}{5-2}
	\cellcolor{red}{5-3}
	\cellcolor{red}{5-4}
	\cellcolor{red}{5-5}
	\cellcolor{red}{5-6}
	\cellcolor{red}{5-7}
	\cellcolor{red}{5-8}
	\cellcolor{red}{5-9}
	\cellcolor{cyan}{5-10}
	\cellcolor{cyan}{5-11}
	\cellcolor{cyan}{5-12}
	\cellcolor{green}{5-13}
	\cellcolor{cyan}{6-10}
	\cellcolor{cyan}{6-11}
	\cellcolor{cyan}{6-12}
	\cellcolor{green}{6-13}
\Body
	& & & & & 1 & 2 & ... & r & r + 1 & ... & b - 1 & b \\
	a - l & a - l + 1 & ... & a - 1 & a & a + 1 & ... & a + r - 1 & a + r & a + r + 1 & ... & a + b - 1 & a + b \\
	2 a - l & ... & ... & ... & 2 a & ... & ... & ... & 2 a + r & ... & ... & ... & 2 a + b \\
	... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... \\
	... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... & ... \\
	p a - l & ... & ... & ... & p a & ... & ... & ... & p a + r & ... & ... & ... & p a + b \\
\end{NiceArray}
\]

\resetarraystretch
\setlength{\arraycolsep}{6pt}

@@

\begin{itemize}
	\item \underline{\(\lambda(x+y) = \lambda(x) + \lambda(y)\)} \\
	In this case, we use the sum-free property of the coloring of \([\![1,p]\!]\) (in block \(\mathcal{C}\), we only 
	use colors \(1, ..., n\)).
	\item \underline{\(\lambda(x+y) \neq \lambda(x) + \lambda(y)\)} \\
	In this case, we do not have information about the color of \(\lambda(x+y)\). Thereby, we want to have 
	\(x+y \in \mathcal{C}\). A simple solution is to limit the horizontal movement, that is if the sum changes line 
	(that is its line number is different from \(\lambda(x) + \lambda(y)\)), not to move too far from \((\lambda(x) 
	+ \lambda(y)) a\) so that the sum stays in \(\mathcal{C}\). There, the maximal displacement to the left (resp. 
	to the right) is \(2l\) (resp. \(2r\)). Not crossing entirely \(\mathcal{C}\) by going to the left is then expressed 
	as \(-2l > -a + r\). Likewise, not going to far to the right is expressed as \(2r < a - l\). It can then be written 
	as \(\max(l, r) \leqslant w\).
\end{itemize}

@@

\qquad In this subsection, we introduce WSF-templates and prove calculative results for the general construction 
theorem on templates for weak Schur numbers.

\begin{definition}
Let \((a,b) \in (\mathbb{N}^*)^2\) such that \(a>b\). Define
\[ \pi_{a,b}:x \longmapsto (\operatorname{Id}+a\mathds{1}_{ [\![0,b]\!]})(x \mod a)\]
\end{definition}

If there is no confusion on the \(a\) and \(b\) to use, \(\pi_{a, b}\) is denoted by \(\pi\). Notice that 
\({\forall x \in \mathbb{Z}, \pi(x) = x \mod a}\) and \({\forall x \in [\![b + 1, a + b]\!], b + 1 \leqslant \pi(x) \leqslant a + b}\).

\begin{sloppypar}
\(\pi\) is the projection on the first line mentioned in the intuitive explanation. 
\end{sloppypar}

\begin{proposition}
\label{prop1}
\[
\forall x \in [\![b + 1, a + b]\!], \pi(x) = x
\]
\end{proposition}

\begin{proof}[\textsc{Proof.}]
\begin{sloppypar}
Let \(x \in [\![b + 1, a + b]\!]\). If \(x < a\) then \(x \mod a = x \notin [\![1, b]\!]\). Hence \(\pi(x) = x\). 
Otherwise, \({x \mod a = x - a \in [\![1, b]\!]}\). Hence \({\pi(x) = x - a + a = x}\).
\end{sloppypar}
\end{proof}

\begin{proposition}
\label{prop2}
Let x \(\in [\![1,b]\!]\) and \(y \in \mathbb{N}^*\). Then 
\[
\pi(x+\pi(y)) = \pi(x+y)
\]
\end{proposition}

\begin{proof}[\textsc{Proof.}]
It is a direct consequence of \(pi(x) = x \mod a\). \\
\end{proof}

\begin{proposition}
\label{prop3}
Let x \(\in [\![1,b]\!]\) and \(y \in \mathbb{N}^*\) such that \(x+\pi(y) \leqslant a+b\). Then 
\[
\pi(x+y)=x+\pi(y)
\]
\end{proposition}

\begin{proof}[\textsc{Proof.}]
\(\pi(y) \geqslant b + 1\) and thus \(b + 1 \leqslant x + \pi(y) \leqslant a + b\).
\[
\begin{array}{l l l l}
	\pi(x + y) & = & \pi(x + \pi(y)) & proposition \ref{prop2} \\
	 & = & x + \pi(y) & proposition \ref{prop1}
\end{array}
\]
\end{proof}

\begin{proposition}
\label{prop4}
Let \((x,y)\in (\mathbb{N}^*)^2\). Then 
\[
\pi(\pi(x)+\pi(y))=\pi(x+y)
\]
\end{proposition}

\begin{proof}[\textsc{Proof.}]
It is a direct consequence of \(\pi(x) = x \mod a\). \\
\end{proof}

\begin{definition}
Let \((a,b) \in (\mathbb{N}^*)^2\) such that \(a>b\). Define
\[ \lambda_{a,b}:x \longmapsto 1+ \left\lfloor\dfrac{x-b-1}{a}\right\rfloor\]
\end{definition}

@@

\begin{proposition}
\label{prop5}
Let \(x\in \mathbb{N}^*\). Then 
\[
x=a\lambda(x)+\pi(x)-a
\]
\end{proposition}

@@

\noindent  if \(x \mod a>b\) then \(a\lambda(x)+\pi(x)-a=a\left\lfloor\dfrac{x}{a}\right\rfloor+x \mod a=x\) \\
\vspace{1mm}
\noindent if \(x \mod a \leqslant b\) then
\(a\lambda(x)+\pi(x)-a=a \left( \left \lfloor \dfrac{x}{a} \right \rfloor - 1 \right)+x \mod a +a=x\) \\
\end{proof}

\begin{proposition}
\label{prop6}
Let \(x, y \in \mathbb{Z}\) such that \(\lambda(x + y) = \lambda(y)\). Then 
\[
\pi(x + y) = x + \pi(y)
\]
\end{proposition}

\begin{proof}[\textsc{Proof.}]
By applying twice proposition \ref{prop5}, \(a \lambda(x + y) + \pi(x + y) - a = x + y = x + a \lambda(y) + \pi(y) - a\). 
The result is then obtained by symplying the equality. \\
\end{proof}

\begin{proposition}
\label{prop7}
Let \(x, y \in \mathbb{Z}\) such that \(\pi(x) + \pi(y) \in [\![a + b + 1, 2 a + b]\!]\). Then 
\[
\lambda(x + y) = \lambda(x) + \lambda(y)
\]
\end{proposition}

\begin{proof}[\textsc{Proof.}]
By proposition \ref{prop5}, \(x + y = a (\lambda(x) + \lambda(y)) + \pi(x) + \pi(y) - 2 a\). Then
\[
\begin{array}{l l l}
	\lambda(x + y) & = & \left \lfloor \dfrac{x + y - b - 1}{a} \right \rfloor + 1 \vspace{1mm} \\
	 & = & \left \lfloor \dfrac{a (\lambda(x) + \lambda(y)) + \pi(x) + \pi(y) - 2 a - b - 1}{a} \right \rfloor + 1 \vspace{1mm} \\
	 & = & \lambda(x) + \lambda(y) - 1 \left \lfloor \dfrac{\pi(x) + \pi(y) - b - 1}{a}\right\rfloor \\
	 & = & \lambda(x)+\lambda(y) -1 +1 ~~~~\text{since } \pi(x) + \pi(y) \in [\![a + b + 1, 2 a + b]\!] \\
	 & =& \lambda(x)+\lambda(y)
\end{array}
\]
\end{proof}

\begin{definition}
Let \( (a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\). Let \((A_1,...,A_n)\) a partition of  \([\![1, a + b]\!]\).
This partition is said to be a \(b\)-weakly-sum-free template (\(b\)-WSF-template) with width \(a\) and \(n\) colors if

\begin{itemize}
\item \(\forall i \in [\![1, n]\!], A_i\) is weakly-sum-free
\item \(\forall i \in [\![1, n]\!], A_i\backslash [\![1, b]\!]\) is sum-free
\item For \(A_n\) (the special subset)
	\[
	\forall (x,y) \in A_n^2, \,x+y>b+2a \implies x+y-2a\notin A_n
	\]
\item For the others subsets
	\[
	\forall i \in [\![0,n-1]\!], \, \forall(x,y) \in A_i^2, \, x+y>a+b \implies \pi(x+y) \notin A_i
	\]
\end{itemize}
\end{definition}

@@

\begin{definition}
Let \( (n,b) \in (\mathbb{N}^*)^2\). If there exists \(a\) such that there exists a \(b\)-WSF-template with width \(a\) 
and \(n\) colors, define
\[
\WS_b^+(n)= \max \{a \in \mathbb{N}^*/ \text{there exists a } b \text{-WSF-template with width } a \text{ and } n \text{ colors} \}
\]
If no such \(a\) exists, set \(\WS_b^+(n) = 0\).
\end{definition}

@@

\begin{proof}[\textsc{Proof.}]
The lower bound comes from the analogous of Abott and Hanson's construction for weak Schur numbers.
The upper bound comes from the fact that a WSF-template with width \(a\) and \(n\) colors contains a partition of
\([\![1, a]\!]\) into \(n\) sum-free subsets. \\
\end{proof}

@@

\begin{theorem}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\) and \( (p,k) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,p]\!]\) 
into \(k\) sum-free subsets and a \(b\)-WSF-template \((A_1,...,A_{n+1})\) with width \(a\) and \(n+1\) colors, 
then there exists a partition of \([\![1, p a + b]\!]\) into \(k+n\) weakly sum-free subsets.
\end{theorem}

In particular, by setting \(p = S(k)\) and \(a = \WS^+(n + 1)\) in the last theorem, the next corollary follows.

\begin{corollary}
Let \(n,k \in \mathbb{N}^*\) and set \( b_{max} = \max \{b\in \mathbb{N}^*/ \WS_b^+(n+1) = \WS^+(n+1)\}\).
Then
\[
\WS(n+k) \geqslant S(k) \WS^+(n+1)+b_{max}
\]
\end{corollary}

\begin{remark}
In the SF-template construction for Schur numbers, the additive constant comes from the fact that the special color does
not necessarily appear right at the begining of the repeating pattern. Likewise, \(b_{max}\) can actually be replaced by
\[
\max_{b \in \mathbb{N}^*} \left\{\min (A_{n+1} \backslash [\![1, b ]\!]) - 1~|~ \WS_b^+(n+1) = \WS^+(n+1) \right\}
\]
\end{remark}

\begin{proof}[\textsc{Proof.}]
\begin{sloppypar}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) and \((p,k) \in (\mathbb{N}^*)^2\). Denote by \(g\) the coloring 
associated to the \(b\)-WSF-template  and \(h\) the one associated to the sum-free partition of \([\![1,p]\!]\); where 
\({g : [\![1, a + b]\!] \longrightarrow [\![1,n+1]\!]}\) and \({h : [\![1, p]\!]  \longrightarrow [\![1, k]\!]}\). Moreover, assume 
that the sum-free coloring of \([\![1, p]\!]\) is ordered.
\end{sloppypar}

\par
NB: In the following 5 predicates, the conditions \(x + y \leqslant p\)  and \(x + y \leqslant a + b\) are omitted for readability.
\par
The (weakly) sum-free conditions are expressed as 
\[
\forall (x,y) \in [\![1,a + b]\!]^2, \left\{
\begin{array}{l}
	g(x) = g(y) \\
	x \neq y
\end{array}
\right. \Longrightarrow g(x+y) \neq g(x)
\]
\[
\forall (x,y) \in [\![b+1,a + b]\!]^2, g(x) = g(y) \Longrightarrow g(x+y) \neq g(x)
\]
\[
\forall (x,y) \in [\![1,p]\!]^2, h(x) = h(y) \Longrightarrow h(x+y) \neq h(x)
\]

The additionnal constraints for the WSF-template are
\[
\forall (x,y) \in [\![1,a + b]\!]^2, \left\{
\begin{array}{l}
	g(x) = g(y) \leqslant n \\
	x + y > a + b
\end{array}
\right. \Longrightarrow g(\pi(x+y)) \neq g(x)
\]
\[
\forall (x,y) \in [\![1,a + b]\!]^2, \left\{
\begin{array}{l}
	g(x) = g(y) = n + 1 \\
	x + y > 2 a + b
\end{array}
\right. \Longrightarrow g(x+y - 2 a) \neq g(x)
\]

Here, we consider \(\pi\) defined in the previous subsection as an application defined on \([\![b + 1, p a + b]\!]\). 
Split \([\![1, p a + b]\!]\) into 3 subsets.

\begin{itemize}
	\item \(\mathcal{T} = [\![1, b]\!]\)
	\item \(\mathcal{C} = \pi^{-1}(g^{-1}([\![1, n]\!]))\)
	\item \(\mathcal{R} = \pi^{-1}(g^{-1}(\{n + 1\}))\)
\end{itemize}

Define a new coloring \(f\) as follows
\[
\begin{array}{c c c l}
	h : & [\![1, p a + b]\!] & \longrightarrow & [\![1,n+k]\!] \\
	& x & \longmapsto & 
	\left\{ \begin{array}{l l}
		g(x) & \text{if}~x \in \mathcal{T} \\
		g(\pi(x)) & \text{if}~x \in \mathcal{C} \\
		n + h(\lambda(x)) & \text{if}~x \in \mathcal{R}
	\end{array} \right.
\end{array}
\]

\(f\) is well defined since \((\mathcal{T}, \mathcal{C}, \mathcal{R})\) is a partition of  \([\![1, p a + b]\!]\). 
We now prove that \(f\) is a weakly sum-free coloring. Let \(x,y \in [\![1, p a + b]\!]\) such that \(x \neq y\), 
\(h(x) = h(y)\) and \(x+y \leqslant p a+ b\). We claim that \(h(x+y) \neq h(x)\). Distinguish between six cases 
according to the subsets \((\mathcal{T}, \mathcal{C}, \mathcal{R})\) to which \(x\) and \(y\) belong. It is 
sufficient to check only six cases out of nine since \(x\) and \(y\) play symmetric roles. \\
